---
title: "Multinomial Logistic Regression"
output: html_document
date: "2024-12-22"
---


- Selected covariates based on an unadjusted analysis of JUSTICE variables stratified by sleep phenotype (see latent_class_analysis.Rmd file) 
- See Table 3 here: https://www.ahajournals.org/doi/10.1161/JAHA.124.035683?url_ver=Z39.88-2003&rfr_id=ori:rid:crossref.org&rfr_dat=cr_pub%20%200pubme

```{r, include=F}
knitr::opts_chunk$set(echo = F, include=T, message=F, warning=F)


check_cat <- function(var_name) {
  
  # Show the current reference category (the first level)
  ref <- levels(sleep_justice[[var_name]])[1]
  
  cat("Current reference category:", ref, "\n")
  
  # Show the counts of each category
  category_counts <- table(sleep_justice[[var_name]], exclude = F)
  print(category_counts)
  
  # Check if any category has very low counts (less than 10)
  low_count_categories <- category_counts[category_counts < 10]
  if (length(low_count_categories) > 0) {
    cat("\nWarning: Categories with low counts (< 10):\n")
    print(low_count_categories)
  }
}
 
check_cat2 <- function(df, cat) {
  # Ensure the input column exists
  if (!cat %in% colnames(df)) {
    stop("The specified categorical variable does not exist in the dataframe.")
  }
  if (!"lc4_class" %in% colnames(df)) {
    stop("The outcome variable 'lc4_class' is not found in the dataframe.")
  }

  # Create a contingency table of the categorical variable against the outcome
  contingency_table <- table(df[[cat]], df$lc4_class)

  # Identify levels with fewer than 10 values in any class of lc4_class
  problematic_levels <- apply(contingency_table, 1, function(row) any(row < 10))

  # Output the problematic levels
  if (any(problematic_levels)) {
    cat("Levels with fewer than 10 values in any class of lc4_class:\n")
    print(contingency_table[problematic_levels, , drop = FALSE])
  } else {
    cat("No levels with fewer than 10 values found.\n")
  }
}
 
  
 impute_cat <- function(data, cat_vars) {
  # Ensure all categorical variables with missing values have 'Unknown' as a level and replace NAs
  data <- data %>%
    mutate(across(all_of(cat_vars), ~{
      if (any(is.na(.))) {  # Only process variables with missing values
        # Add "Unknown" as a level if it's not already a factor level
        if (!"Unknown" %in% levels(.)) {
          . <- factor(., levels = c(levels(.), "Unknown"))
        }
        # Replace NA with "Unknown"
        replace(., is.na(.), "Unknown")
      } else {
        .  # If no NAs, leave the variable as is
      }
    }))
  
  return(data)
}

# Define the function
create_histogram <- function(data, x_var) {
  ggplot(data, aes_string(x = x_var)) +  # Use aes_string to reference the variable dynamically
    geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +  # Histogram
    geom_vline(aes_string(xintercept = paste("median(", x_var, ", na.rm = TRUE)", sep = "")), 
               color = "red", linetype = "dashed", size = 1) +  # Add vertical line at the median
    labs(title = paste("Histogram of", x_var), x = x_var, y = "Frequency") +  # Dynamic title and x-axis label
    theme_minimal()
}

summarize_model <- function(model, model_name) {
  cat(paste("Number of covariates in", model_name, ":", length(model$coefnames)), "\n")
  cat(paste(model$coefnames, collapse = "\n"), "\n")
}


get_stars = function(p){
    case_when(p <= 0.001 ~ "***",
        p <= 0.01 ~ "**",
        p <= 0.05 ~ "*",
        p <= 0.1 ~ ".",
        TRUE ~ " ")
        }


# Define a function to calculate p-values, exponentiated coefficients, and confidence intervals
get_model_results <- function(model) {
  z <- summary(model)$coefficients / summary(model)$standard.errors
  p <- (1 - pnorm(abs(z), 0, 1)) * 2
  coef_exp <- exp(coef(model))
  conf_int <- exp(confint(model, level = 0.95))
  
  list(p_values = p, exp_coef = coef_exp, conf_int = conf_int)
}

# Function to get results_df for "lesion" terms
get_results_df <- function(model) {
  tidy(model, exponentiate = TRUE, conf.int = TRUE) %>%
    mutate(stars = get_stars(p.value)) 
}



# Define a function to create the multinomial logistic regression plot
plot_or <- function(model, model_data, title) {
  
  # Extract coefficients and confidence intervals
  tidy_results <- broom::tidy(model, conf.int = TRUE)
  
  # Convert estimates to odds ratios (exponentiate coefficients)
  tidy_results <- tidy_results %>%
    mutate(estimate = exp(estimate),
           conf.low = exp(conf.low),
           conf.high = exp(conf.high)) %>%
    filter(term != "(Intercept)") %>%
    rename(Class = y.level)

  # Define significance levels
  tidy_results <- tidy_results %>%
    mutate(significance = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      TRUE            ~ ""
    ))

  # Define variable label mappings (ensuring alignment with the table)
  variable_labels <- c(
    "paroleYes" = "Current parole (Yes)",
    "Gen_healthFair/poor" = "General Health (Fair/Poor)",
    "ptsd_score" = "Total PTSD Score",
    "Adv_score" = "Total Lifetime Adversity Score",
    "PHQ9_score" = "Total PHQ9 Score",
    "GSE_score" = "Self-Efficacy Score",
    "Pss_score" = "Perceived Stress Scale Score",
    "Discrim_score" = "Perceived Discrimination Score",
    "EducationHS" = "Education (HS)",
    "EducationLessHS" = "Education (Less than HS)",
    "RaceHispanic" = "Race (Hispanic)",
    "Age" = "Age at First Visit"
  )

  # Apply labels
  tidy_results$term <- factor(tidy_results$term, levels = names(variable_labels), labels = variable_labels)

  # Define custom color mapping
  custom_colors <- c("Moderate-quality controlled sleep\n(Class 2)" = "#7CAE00",
                     "Poor-quality limited control sleep with OSA risk\n(Class 3)" = "#00BFC4",
                     "Severely impaired uncontrolled sleep with insomnia\n(Class 4)" = "#F8766D")

                           #putting 1-4 for the figure even though the levels names are different 

  # Plot using ggplot2
  ggplot(tidy_results, aes(x = estimate, y = term, color = Class)) +
    geom_point(size = 2) +
    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
    facet_wrap(~ Class, scales = "free") +
    geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
    scale_x_log10(breaks = c(0.01, 0.1, 1, 10, 100), labels = c("0.01", "0.1", "1", "10", "100")) +
    scale_color_manual(values = custom_colors) +
    geom_text(aes(label = significance), nudge_x = 0.2, size = 5, color = "black") +  # Add asterisks
    theme_minimal() +
    labs(title = title,
         x = "Odds Ratio",
         y = "Predictor Variable") +
    theme(legend.position = "none")
}


```




```{r setup, include=FALSE}
library(nnet)
library(tidyverse)
library(naniar)
library(table1)
library(MASS)
library(broom)
library(forcats)
library(gtsummary)

load("../data/clean/sleep_justice.Rda")


dd = read_csv("../resources/data_dictionary_justice.csv", show_col_types = F, col_select = c("variable", "type"))


my_vars <- c("participant_id", "lc4_class", "Age", "Race", "self_gender", "Education", "Civil_status","Income_4cat",
                   "Discrim_score", "Pss_score", "GSE_score", "PHQ9_score", "Adv_score", "ptsd_score","Gen_health",
                   "Residence",
                   "parole", "probation", "solitary", "Sol_time",
                   "Food_ins",
                   "psqi_sleep_satisfaction", "psqi_sleep_latency", "psqi_sleep_duration", 
                   "psqi_sleep_efficiency", "psqi_sleep_disturbance", "psqi_sleep_alertness", 
                   "bangosa_score_cat", "psqi_insomnia", "psqi_med")



df2 = sleep_justice %>%
    dplyr::select(all_of(my_vars)) %>%
    mutate(
        Residence = relevel(Residence, "Transitional home"),
        Gen_health = relevel(Gen_health, "Good to Excellent")
        )  %>%
    dplyr::select(-self_gender, -starts_with("psqi"), -bangosa_score_cat, -probation, -Food_ins, -Sol_time)


my_vars2 = names(df2)

dd2 = dd %>%
    filter(variable %in% my_vars2)  
           

cat_vars = dd2 %>% filter(type %in% c("category","binary")) %>% pull(variable)
num_vars = dd2 %>% filter(type == "continuous") %>% pull(variable)

```



```{r}
df3 <- df2 %>%
    impute_cat(., cat_vars) %>%
    drop_na() %>%
    mutate(across(all_of(cat_vars), droplevels))
```


```{r, include=F}
set.seed(1)
m1 = multinom(lc4_class ~ . - participant_id, data=df3)
```

```{r, include=F}
m2 <- stepAIC(m1, direction = "both", trace = FALSE)
```


```{r}
paste("Number of covariates in Model 1:", length(m1$coefnames))
paste(m1$coefnames)
paste("Number of covariates in Model 2:", length(m2$coefnames))
paste(m2$coefnames)
```

Need to drop:
Residence: Very low counts of all levels except transitional housing.
Income: Very low counts of all levels except No income
Civil Status: Very low counts of all levels except Single, need to drop
Solitary: Lots of missing, so unknown is coming up as a coefficient

Race: Low counts of NH white and NH other, will need to create a hispanic binary variable
Education: Need to combine HS grad and GED together, Some college/College grad


```{r}

df2 = sleep_justice %>%
    dplyr::select(all_of(my_vars)) %>%
    mutate(
        Residence = relevel(Residence, "Transitional home"),
        Gen_health = relevel(Gen_health, "Good to Excellent"),
        Race = fct_collapse(Race, 
                            NonHispanic = c("Non-hispanic white", "Non-hispanic black","Non-hispanic other"),
                            Hispanic = c("Hispanic")),
        Education = fct_collapse(Education,
                                 College = c("Some college/tech",
                                             "College grad/grad school"),
                                 HS = c("HS grad","GED"),
                                 LessHS = c("Less than HS")),
        Education = relevel(Education, "College"))%>%
    dplyr::select(-self_gender, -starts_with("psqi"), -bangosa_score_cat, -probation, -Food_ins, -Sol_time, -Residence, -Income_4cat, -Civil_status, -solitary)


my_vars2 = names(df2)

dd2 = dd %>%
    filter(variable %in% my_vars2)  
           

cat_vars = dd2 %>% filter(type %in% c("category","binary")) %>% pull(variable)
num_vars = dd2 %>% filter(type == "continuous") %>% pull(variable)



```

```{r}
df3 <- df2 %>%
    impute_cat(., cat_vars) %>%
    drop_na() %>%
    mutate(across(all_of(cat_vars), droplevels))
```

```{r, include=F}
set.seed(1)
m1 = multinom(lc4_class ~ . - participant_id, data=df3)
```

```{r, include=F}
m2 <- stepAIC(m1, direction = "both", trace = FALSE)
```

```{r}
summarize_model(m1, "Model 1")
```
```{r}
summarize_model(m2, "Model 2")
```

Interpretation: After adjusting for confounders, a higher PHQ-9 score and total lifetime adversity score was significantly associated with increased odds of poor sleep/no insomnia (class 3) and worst sleep w/insomnia (class 4) compared to the reference group (class 1). A higher self-efficacy score was associated with moderately good sleep (class 2). 

```{r}
tbl_regression(
  m1, 
  exponentiate = TRUE, 
  include = everything(),
)

```


```{r}
tbl_regression(
  m2, 
  exponentiate = TRUE, 
  include = everything(),
)


```


```{r, fig.width=14, fig.height=6, out.width="100%"}
p2 = plot_or(m1, df3, "Multinomial Logistic Regression Results (Model 1)")
```


```{r}
ggsave("forest_plot.png", plot = p2, dpi = 300, width = 14, height = 6, units = "in")
ggsave("forest_plot.tiff", plot = p2, dpi = 300, width = 14, height = 6, units = "in", compression = "lzw")
```

```{r, fig.width=12, fig.height=6, out.width="100%"}
plot_or(m2, df3, "Multinomial Logistic Regression Results (Model 2)")
```

